---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dlw

<!-- badges: start -->
<!-- badges: end -->

`{dlw}` is an R client of the internal [datalibweb](https://datalibweb2.worldbank.org/) API of the World Bank. Datalibweb is a data system designed to enable users to seamlessly access microdata and documentation in the World Bank. Users can access the most up-to-date and historical versions of harmonized data collections and raw/non-harmonized data for subsequent analysis

## Installation

You can install the development version of dlw from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("worldbank/dlw")
# or
remotes::github_install("worldbank/dlw")
```

## Usage
```{r}
library(dlw)
```

### Token
To begin working with `{dlw}`, you must provide your datalibweb API token using the `dlw_set_token()` function. 

You can obtain or renew your token by visiting the [datalibweb page](https://datalibweb2.worldbank.org/) and following the instructions there. Once you have your token, set it in your R session as follows:

```{r, eval = FALSE}
dlw_set_token("your_token_here")
```

### server catalob
After setting your token, you can download the catalog for the corresponding server using `dlw_server_catalog()`. By default, this downloads the "GMD" catalog. 
The downloaded catalog is saved for the current session in a hidden environment within the `{dlw}` package, allowing for easy and efficient access in subsequent operations.

```{r example}
ctl <- dlw_server_catalog()
```

The `{dlw}` package returns results as `data.table` objects, enabling lightning-fast data manipulation and filtering using concise syntax. For example, you can instantly list all available files for Colombia in 2010 in module "ALL" of the GMD collection with a single line of code:

```{r}
ctl[Country_code == "COL" & Module == "ALL" & Survey_year == 2010, 
  .(FileName, Vermast, Veralt )]
```

### Downloading files

The workhorse function to download data is `dlw_get_data()`. However, it requires several pieces of information that you may not have at hand, such as:

```r
dlw_get_data(
  country_code = "PRY",
  year = 2011L,
  server = "GMD",
  survey = "EPH",
  module = "GPWG",
  filename = "PRY_2011_EPH_V01_M_V03_A_GMD_GPWG.dta",
  collection = "GMD"
)
```

To simplify this process, we have developed a wrapper function that works only for the GMD server: `dlw_get_gmd()`. This function is much easier to use:

```{r}
pry <- dlw_get_gmd(country_code = "PRY", year = 2011, module = "GPWG", vermast = "v01", veralt = "v03")
pry[, weighted.mean(welfare, weight, na.rm = TRUE)]
```

If you are interested in downloading the most recent version of a file, you can simply omit the version arguments:

```{r}
pry <- dlw_get_gmd(country_code = "PRY", year = 2011, module = "GPWG")
pry[, weighted.mean(welfare, weight, na.rm = TRUE)]
```

In this case, the most recent version available will be used (for example, master version v02 and alternative version v01).

