---
title: "dlw"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dlw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dlw)
```

`{dlw}` is an R client of the internal [datalibweb](https://datalibweb2.worldbank.org/) API of the World Bank. Datalibweb is a data system designed to enable users to seamlessly access microdata and documentation in the World Bank. Users can access the most up-to-date and historical versions of harmonized data collections and raw/non-harmonized data for subsequent analysis


## Token
To begin working with `{dlw}`, you must provide your datalibweb API token using the `dlw_set_token()` function. 

You can obtain or renew your token by visiting the [datalibweb page](https://datalibweb2.worldbank.org/) and following the instructions there. Once you have your token, set it in your R session as follows:

```{r, eval = FALSE}
dlw_set_token("your_token_here")
```

[TO COMPLETE]
