---
title: "dlw"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dlw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dlw)
library(data.table)
```

`{dlw}` is an R client of the internal [datalibweb](https://datalibweb2.worldbank.org/) API of the World Bank. Datalibweb is a data system designed to enable users to seamlessly access microdata and documentation in the World Bank. Users can access the most up-to-date and historical versions of harmonized data collections and raw/non-harmonized data for subsequent analysis

## Token

To begin working with `{dlw}`, you must provide your datalibweb API token using the `dlw_set_token()` function.

You can obtain or renew your token by visiting the [datalibweb page](https://datalibweb2.worldbank.org/) and following the instructions there. Once you have your token, set it in your R session as follows:

```{r, eval = FALSE}
# dlw_set_token("your_token_here")
```

## Server Catalog

The server catalog provides a comprehensive list of all datasets available across all countries in the datalibweb system. You can retrieve the server catalog using the `dlw_server_catalog()` function.

```{r}
server_cat <- dlw_server_catalog()
head(server_cat)
```

### Exploring the Server Catalog

You can filter and explore the server catalog to find datasets of interest.

```{r}
# For examples,
# to find list of all datasets related to HEIS surveys:
heis_datasets <- server_cat[Survey == "HEIS", ]
head(heis_datasets)

# to find list of all datasets for a specific module:
module_gpwg <- server_cat[Module == "GPWG", ]
head(module_gpwg)

# to find list of all datasets for a specific year:
year_2023 <- server_cat[Year == 2023, ]
head(year_2023)

# to find list of all datasets for a specific country and year:
country_year <- server_cat[Country == "PRY" & Year == 2020, ]
head(country_year)
```

## Country Catalog

The country catalog provides detailed metadata for datasets specific to a single country. You can retrieve the country catalog using the `dlw_country_catalog()` function by specifying the desired country code.

```{r}
country_pry <- dlw_country_catalog(country_code = "PRY")
head(country_pry)
```

### Exploring the Country Catalog

You can filter and explore the country catalog to find list datasets of interest.
```{r}
# For examples 
# to find list of all datasets related to EPH survey in Paraguay: 
eph_pry <- country_pry[Survey == "EPH", ] 
head(eph_pry)

# to find list of all datasets for a specific year:
year_2020_pry <- country_pry[Year == 2020, ] 
head(year_2020_pry)
```

## Downloading Datasets
Once you have identified datasets of interest from either the server or country catalog, you can download them using the `dlw_get_data()` function.
```{r}
dlw_get_data(
  country_code = "PRY",
  year = 2020L,
  server = "GMD",
  survey = "EPH",
  module = "GPWG",
  filename = "PRY_2020_EPH_V01_M_V03_A_GMD_GPWG.dta",
  collection = "GMD"
)
```

Alternatively, you can use `dlw_get_gmd()`, which is a convenient wrapper around `dlw_get_data()`. To download the most recent version of a dataset, you only need to provide the country code, year, and module.
```{r}
pry20 <- dlw_get_gmd(country_code = "PRY", year = 2020, module = "GPWG")
```

To download a specific version of a dataset, you can provide additional arguments such as `survey`, `vermast`, and `veralt` to `dlw_get_gmd()`.
```{r}
# this downloads version V02 of the EPH survey for Paraguay in 2020 within the GPWG module.
pry20_v2 <- dlw_get_gmd(
  country_code = "PRY",
  year = 2020,
  module = "GPWG",
  survey = "EPH",
  vermast = "V01",
  veralt = "V02"
)
```
