---
title: "Country Catalog"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Country Catalog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_token <- tryCatch(
  {
    dlw::dlw_check_token()
    TRUE
  },
  error = function(e) FALSE
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = has_token
)
```

## Introduction

The **dlw** package provides granular access to country-specific dataset information from the Datalibweb (DLW) API. While `dlw_server_catalog()` gives you a complete list of all datasets across all countries, `dlw_country_catalog()` provides detailed, country-level metadata for a specific nation.

This vignette covers:

- Understanding the difference between server and country catalogs
- Retrieving country-specific catalog information
- Filtering and exploring country-level datasets
- Leveraging caching for performance

## Prerequisites

Before using `dlw_country_catalog()`, ensure you have:

1. Installed the **dlw** package
2. Set your API token using `dlw_set_token()`

```{r setup}
# Install dlw (if not already installed)
# devtools::install_github("worldbank/dlw")

# devtools::load_all(".")
library(dlw)
library(data.table)

# Set your API token
# dlw_set_token("your_api_token_here")
```

## Server Catalog vs. Country Catalog

Before diving into `dlw_country_catalog()`, understand the difference:

| Aspect | Server Catalog | Country Catalog |
|--------|---|---|
| **Scope** | All datasets across all countries | Datasets for a single country |
| **Function** | `dlw_server_catalog()` | `dlw_country_catalog()` |
| **Detail Level** | More information | Country-specific info |
| **Use Case** | Overview, across countries | Deep dive into a specific country |
| **Size** | Larger dataset | Smaller, focused dataset |

**When to use each:**

Use country catalog function to explore all available data for a particular country, e.g., Colombia
```{r}
country_cat <- dlw_country_catalog("COL")
head(country_cat)
```

Use server catalog function to get the list of datasets available across all countries.
```{r comparison}
server_cat <- dlw_server_catalog()
head(server_cat)

```

## Basic Usage: Retrieve Country Catalog

The simplest way to get a country catalog is:

```{r basic-usage}
# Retrieve the public catalog for Colombia
catalog_col <- dlw_country_catalog(country_code = "COL")

# View the first few rows
head(catalog_col)

# Check the dimensions
dim(catalog_col)
```

This returns a `data.table` with detailed metadata about all datasets available for Colombia.

### Understanding the Catalog Structure

```{r catalog-structure}
# View column names
colnames(catalog_col)

# Check the structure
str(catalog_col)

```

Key columns typically include:

| Column | Type | Description |
|--------|------|-------------|
| **ServerAlias** | character | Server name |
| **Country** | character | ISO3 country code |
| **Year** | integer | Year of the survey |
| **FilePath** | character | Full file path on the server |
| **Ext** | character | File extension |
| **FileSize** | number | File size |
| **Timestamp** | character | Date of data compilation |
| **Checksum** | character | hash |

## Exploring Country Catalog

### Get Summary Statistics

```{r summary-stats}
# How many surveys for this country?
unique_surveys <- catalog_col[, .(count = .N), by = Survey]
print(unique_surveys)

# Year range of surveys
year_range <- catalog_col[, .(
  min_year = min(Year, na.rm = TRUE),
  max_year = max(Year, na.rm = TRUE),
  num_years = uniqueN(Year)
)]
print(year_range)

```

### Filter by Survey Type

```{r filter-survey}
# Find all GEIH surveys
geih_surveys <- catalog_col[Survey == "GEIH"]
head(geih_surveys)

# Find all surveys from 2010 onwards
recent_surveys <- catalog_col[Year >= 2010]
nrow(recent_surveys)

# Find HBS surveys
hbs_surveys <- catalog_col[Survey == "HBS"]
if (nrow(hbs_surveys) > 0) {
  print(hbs_surveys)
} else {
  cat("No HBS surveys found for this country\n")
}
```

### Find Latest Versions

```{r latest-versions}
# For each survey, find the latest year
latest_by_survey <- catalog_col[, .SD[which.max(Year)], by = Survey]
print(latest_by_survey[, .(Survey, Year)])

# Find the absolute latest survey
latest_overall <- catalog_col[order(-Year)][1]
print(latest_overall)
```

## Working with Multiple Countries

Compare catalogs across different countries:

```{r multiple-countries}
# Get catalogs for multiple countries
countries <- c("COL", "BRA", "IND", "MEX")

# Create a list to store catalogs
catalogs <- lapply(countries, dlw_country_catalog)
names(catalogs) <- countries

# Compare dataset counts
dataset_counts <- data.table(
  Country = countries,
  Datasets = sapply(catalogs, nrow)
)
print(dataset_counts)

# Find common surveys across countries
surveys_by_country <- lapply(catalogs, function(x) unique(x$Survey))
common_surveys <- Reduce(intersect, surveys_by_country)
cat("Surveys available in all countries:", paste(common_surveys, collapse = ", "), "\n")

# Find unique surveys per country
for (country in countries) {
  unique_surveys <- setdiff(
    unique(catalogs[[country]]$Survey),
    common_surveys
  )
  if (length(unique_surveys) > 0) {
    cat(country, "unique surveys:", paste(unique_surveys, collapse = ", "), "\n")
  }
}
```

## Understanding Caching

By default, `dlw_country_catalog()` caches country catalogs in memory during your R session. This improves performance for repeated calls:

```{r caching}
# First call: queries the API (slower)
system.time({
  cat1 <- dlw_country_catalog("COL")
})

# Second call: uses cache (much faster)
system.time({
  cat2 <- dlw_country_catalog("COL")
})

# Verify they're identical
identical(cat1, cat2)

# Force a fresh API call
system.time({
  cat_fresh <- dlw_country_catalog("COL", force = TRUE)
})
```

### When to Use `force = TRUE`

- **Development/testing**: When you expect the server catalog to have changed
- **Production workflows**: Occasionally refresh to catch new datasets
- **Debugging**: When troubleshooting connectivity issues
- **Multi-session workflows**: When data may have been updated

```{r refresh-catalog}
# Refresh the catalog from the server
catalog_latest <- dlw_country_catalog(
  country_code = "COL",
  force = TRUE,
  verbose = TRUE
)
```

## Troubleshooting

### No Data Returned

If `dlw_country_catalog()` returns an empty result:

```{r troubleshooting-empty}
# Check if the country code is valid
catalog <- dlw_country_catalog("COL")
if (nrow(catalog) == 0) {
  cat("No datasets found for this country\n")
  
  # Try refreshing
  catalog <- dlw_country_catalog("COL", force = TRUE)
}
```

### Authentication Errors

If you get authentication errors:

```{r troubleshooting-auth}
# Verify token is set
# token <- dlw_get_token()
# if (is.null(token) || token == "") {
  # dlw_set_token("your_api_token_here")
# }

```

### API Connection Issues

If you cannot reach the server:

```{r troubleshooting-connection}
# Force refresh to test connection
tryCatch({
  catalog <- dlw_country_catalog("COL", force = TRUE, verbose = TRUE)
  cat("Connection successful\n")
}, error = function(e) {
  cat("Connection failed:", e$message, "\n")
})
```

## Related Functions

- `dlw_server_catalog()` — List of all datasets across all countries
- `dlw_server_inventory()` — Filter server catalog by multiple criteria
- `dlw_get_data()` — Download data
- `dlw_set_token()` — Set API authentication token

For more information, use:

```r
?dlw_country_catalog
?dlw_get_data
?dlw_server_catalog
```
